Untwisting the Tetrahelix (v0.6)

### 2017-02-24 doc/TetrahelixUntwist.tex
major read through

\begin{abstract}
  The Boerdijk--Coxeter helix (BC helix, or tetrahelix) is a
  face-to-face stack of regular tetrahedra.  Considering the edges of
  these tetrahedra, the resulting structure is attractive and
  inherently rigid, and therefore interesting to architects,
  mechanical engineers, and robotocists.  A formula that matches the
  visually apparent helices forming the outer rails of the tetrahelix
  is derived which is convenient for designers.  This formula defines
  the vertices of tetrahelices of varying radius, pitch, and
  curvature, with the BC helix as a special case.  An addtional
  special case is of that of $0$ curvature or rail angle, which
  generates a \emph{tetrabeam}.  A particular choice of paramaters
  defines a novel object, the minimax member length-difference
  tetrabeam, the \emph{equitetrabeam}.  A formula for computing the
  radius and z-axis travel that gives an optimal minimax tetrahelix
  for any rail angle between $0$ and $\rho_{bc}$ is given, thus
  defining a continuum of tetrahelices containing the equitetrabeam
  and BC helix. Numerically finding $\rho$ from the equation for
  pitch allows optimal tetrahelices of any pitch to be designed.
  An interactive tool for such design and experimentation is provided.
  Utility and use for static and variable geometry
  truss/space frame design and robotics based on the are discussed.
\end{abstract}

@@

The Boerdijk--Coxeter helix\cite{coxeter1985simplicial} (BC helix), is
a face-to-face stack of tetrahedra that winds about a straight axis.
Because architects, structural engineers, and robotocists are inspired
by and follow such mathematical models but can build structures and
machines of differing or even dynamically changing length, it is
useful to develop the mathematics of structure formed from tetrahedra
where we relax regularity.  The vertices of the tetrahedra lie upon
three helices about the central axis.  The
Tetrobot/Glussbot\cite{TetrobotBook} project uses the regularity of
this geometry to make a tentacle-like robot that can crawl like a slug
or mollusc.  The Tetrobot concept is to use mechanical members, called
actuators, which can change their length, connected by special joints,
called the Song-Kwon-Kim\cite{song2003spherical} or turret joint,
which allow many members to come to a single point.  Such machines can
follow purely regular mathematical models such as the Boerdijk–-Coxeter
helix or the Octet Truss\cite{richard1961synergetic}.

@@

BC helix does not rest on a plane in a simple way. It is convenient to
be able to ``untwist'' it and form a tetrahelix space frame that has a
flat planar surface. By making length changes in a certain way, we can
untwist a tetrahelix to form a \emph{tetrabeam} which has planar faces
and has, for example, an equilateral triangular profile. This paper
develops the equations needed to untwist the tetrahelix. All math
deveoped here is available in Javascript and demonstrated by an interactive
design website\cite{readtetrahelix}, from which Figures \ref{fig:closeup} and
\ref{fig:series} are taken.

@@

%% Note: This is not rendering figure two for some reason.
In Figure
\ref{fig:series}, the closest helix is the BC helix.

@@

If you can choose member lengths, you can form a linear combination of
the equitetrabream lengths and the completely regular lengths of the
tetrahelix, thereby choosing the torsion.  If you are designing a
space frame, this is a static design choice, in a robot, it is a
dynamic choice that can be used to twist the robot and/or exert
torsion on the environment.

Ideally we would have a simple formula for defining the nodes based on
any torsion we choose.  Unfortunately, it is not obvious that a linear
combination of lengths produces a simple formula.  It is a goal of
this paper to relate these two approaches to generating a tetrahelix
continuum.

@@

From R.W. Gray's site\cite{graytetrahelix}, repeating formula by Coxeter\cite{coxeter1985simplicial}, for a counter-clockwise BC Helix:
\begin{equation}
  \label{graycoxeter}
V(n) =
\left [
  \begin{tabular}{c}
   $ r_{bc} \cos(n \theta) $\\
   $ r_{bc} \sin(n \theta) $\\
   $ n h_{bc}  $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
 $ r_{bc} = \frac{3\sqrt{3}}{10} $\\
 $ h_{bc} = 1/\sqrt{10} $ \\
 $ \theta = \arccos(-2/3) $ \\
  \end{tabular}      
\end{equation}
where $n$ represents each integer numbered node in succession on every colored rail.

@@

Such a helix can be written:
\begin{equation}
H_{BCcolored}(n,c) =
\left [
  \begin{tabular}{c}
   $ r_{bc}  \cos((3 \theta - 2 \pi)n + c  \theta $\\
   $ r_{bc} \sin((3 \theta - 2 \pi)n + c  \theta $\\
   $ (n + c/3) 3  h_{bc} $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
 $ r_{bc} = \frac{3\sqrt{3}}{10} $\\
 $ h_{bc} = 1/\sqrt{10} $ \\
 $ \theta = \arccos(-2/3) $ \\
  \end{tabular}      
\end{equation}

@@

The quantity $ (3 \theta - 2 \pi) \approx 35.43 \degree $, and is the angular shift between $V(n*3+c)=H_{BCcolored}(n,c)$ and
$V((n+1)*3+c)=H_{BCcolored}(n+1,c)x$.
This quantity appears so often that we call it the ``rail angle rho''. For the BC helix, $\rho_{bc} = (3 \theta - 2 \pi)$.

@@

Since:
\[ \frac{2 \pi}{\rho_{bc}} \approx 10.16
\]
We can see that there are approximately $10.16$ red, blue or yellow tetrahedra on one rail in a single revolution.
The pitch of the Boerdijk--Coxeter helix of edge length $1$ is the length of three tetrahedra times this number:
\begin{align*}
  &= \frac{3 \cdot h_{bc} 2 \pi }{\rho_{bc}} \\
  &= \frac{3  \sqrt{\frac{2}{5}}  \pi}{\rho_{bc}} \\
  &\approx 9.64 \\
\end{align*}
The pitch is less than the number of tetrahedra because the tetrahedra
are not lined up perfectly.  It is a famous and interesting result
that the pitch is irrational, a BC helix never has two tetrahedra at
precisely the same orientation around the $z$-axis. However, this is
inconvenient to designers, who might prefer a rational pitch. For
example, a slight irregularity that led to a pitch of precisely 10
tetrahedra in one revolution would allow an architect to design a
column having a basis and a capital in the same relation to the
tetrahedra they touch.

A BC helix has the useful property that every member is precisely the
same length. If we relax this, so that the tetrahedra it comprises are
not perfectly regular, then we can twist and curve the tetrahelix into
a variety of shapes. This is useful to the mechanical engineer or
robotocist because the structure remains an inherently rigid,
omni-triangulated space frame, which may be expected to be at least
somewhat mechanically strong.

@@

We use the term \emph{tetrahelix} to mean any structure made of
vertices and edges which is isomorphic to the BC helix, in which the
vertices lie on three helices, no matter what lengths the edges take
on. One could consider various definitions of optimality for a
tetrahelix, but the must useful to an enginner is to minimize the
maximum difference between any two edges. We call a tetrahelix
\emph{member-length minimax optimal} if there is not tetrahelix of the
same radius and pitch with a smaller maximum edge length difference.

@@

 Now now that we have coincident axes, the same pitch, we can go on to
 the hard proof about where they occur.

@@

 \begin{theorem}
   \label{eventhirds}
   A tetrahelix of a given radius and height $h$ in which all nodes are evenly spaced at $h/3$ intervals on the $z$ axis is optimal.
     Any one tetrahedron in a tetrahelix has $1$ rail edge, $2$ one-hop edges connected to the rail and $2$ two-hop edges connected to the rail.
  The edge opposite of the rail edge is a one-hop edge.
  \end{theorem}

 \begin{proof}

   In principle in any tetrahelix where the three helices have any
   displacement along the $z$ axis there are 9 distinct edge classes:
   3 rail edges, a one-hop length between each of 3 rails, and two hop
   length between each of three rails, where the two-hop length is a
   maximum length between rails (which could equal the one-hop
   length.) However, we have already shown the pitches and the rail
   lengths are equal in any optimal tetrahelix.

    Consider the tetrahelix in which the vertices are evenly spaced at
    $h/3$ intervals on the $z$ axis. Every edge is either a rail edge,
    or it makes one hop, or it makes two hops. All of the one-hop
    edges are equal length.  All of the two-hop edges are equal
    length.

@@

Note that based on \ref{eventhirds}, we are justified in classifying edge lengths as \emph{rail},\emph{one-hop}, or
\emph{two-hops}. The one-hop edges are the edges between closest on the $z$-axis, and the two-hop edges are those that hop over a vertex.

By \ref{eventhirds} every optimal tetrahelix has vertices lying on helices expressible in the form:
\[
V_{optimal}(n,c) =
\left [
  \begin{tabular}{c}
   $ r \cos(n \alpha +  c 2 \pi /3)$\\
   $ r \sin(n \alpha +  c 2 \pi /3) $\\
   $ \frac{d(n +c / 3)}{3}   $
  \end{tabular}
  \right ],
\text{where:}
\begin{tabular}{c}
  $c \in \{0,1,2\}$
  \end{tabular}      
\]
where we have not yet investigated in the general case the relationships beteween $\alpha$, $r$, and $d$ in this formulation.
However, we understand that when $\alpha = 0$, the helices are degenerate, having curvature of $0$, and
we have the equitetrabeam.

This formulation $V(n,c)$ above is valuable, but obscures the essentially fact that the red, yellow, and blue helices distributed
about the central $z$ axis $120\degree$ from each other.
In order to rewrite this expression with an explicit rotation of $2\pi/3$, we expand 
the expression and seek to isolate the term $c2\pi/3$.
\begin{align*}
  \rho_{bc} n + c \theta  &=   \text{\{we aim for 3 in denominator, so we split...\}} \\
    (3 \theta - 2 \pi)n + (c/3)  (\theta /3)  &=   \text{\{we want $2\pi$ in numerator, so add canceling terms...\}} \\
  (3 \theta - 2 \pi)n + (c/ 3) (3 \theta - 2 \pi  + 2 \pi) &=  \text{\{associate...\}} \\
  (3 \theta - 2 \pi)n + (c/ 3) ((3 \theta - 2 \pi)  + 2 \pi) &=  \text{\{distribute...\}} \\  
  (3 \theta - 2 \pi)n + (c / 3) (3 \theta - 2 \pi)  + c 2 \pi /3 &=  \text{\{definition of $\rho_{bc}$...\}} \\
  \rho_{bc} n + (c / 3) \rho_{bc}  + c 2 \pi /3 &=  \text{\{collect like factors...\}} \\  
  \rho_{bc} (n + c/3)  + c 2 \pi /3  \\
\end{align*}
Now the the term on the left is the only one that depends on the scalar $n$. We use this to a create
a new formulation $H_{BCsymmetric}(n,c) = H_{BCcolored}(n,c)$

@@

\section{Parametrizing Tetrahelices via Rail Angle}

We seek a formula to generate optimal tetrahelices that accepts a
parameter that allows us to choose the tetrahelix conveniently. The
pitch of the helix is an obious choice, but is not defined when the
curvature is $0$, and important special case. The radius or the axial
distance between two nodes on the same rail are obvious choices, but
perhaps the clearest choice is to build formula that takes as its
input the ``rail angle'' $\rho$. We define $\rho$ to be the angle
formed in the X,Y plane $\angle R_i O R_{i+1}$ projecting out the $z$
axis and sighting along the positive $z$ axis. In other words, $\rho$
controls how far a rail edge of a tetrahelix deviates from being
parallel with the axis, or the ``twistiness'' of tetrahelix. We use
the parameter $\chi = 1$ to indicate a chirality or counter-clockwise,
and $\chi = -1$ for clockwise.

Please refer back to Figure \ref{railanglefig}.

@@

In choosing $\rho$, one greatly constrains $r$ and $h$, but does not completely
determine both of them together, so we treat both as parameters.

Rewriting our formulation in terms of $\rho$:
\begin{equation}
H_{general}(\chi,n,c,\rho,d_{\rho},r_{\rho}) =
\left [
  \begin{tabular}{c}
   $ r_{\rho} \cos(\chi \cdot (\rho \kappa + s\cdot c 2 \pi /3)) $\\
   $ r_{\rho}  \sin(\chi \cdot (\rho \kappa + s\cdot c 2 \pi /3)) $\\
   $ d_{\rho} \kappa $
  \end{tabular}
  \right ],
\text{where:}
\begin{tabular}{c}
  $   1 = d_{\rho}^2 + 4 r_{\rho}^2 (\sin{ \rho / 2})^2 $ \\
  $\kappa = n + c/3$ \\
    $\chi \in \{-1,1\}$ \\  
  \end{tabular}      
\end{equation}

@@

\begin{proof}
  This requires a minimax argument.
\end{proof}

By Theoerm \ref{eventhirds}, we can compute the (at most) three edge-lengths of an optimal
tetrahelix by (where $dist$ is the cartesian distance function):
\begin{align*}
  \text{rail} &= dist(H_{general}(n,c,\rho,d_{\rho},r_{\rho}),H_{general}(n+1,c,\rho,d_{\rho}),r_{\rho})) = 1 \\
  \text{one-hop} &= dist(H_{general}(n,c,\rho,d_{\rho},r_{\rho}),H_{general}(n,c+1,\rho,d_{\rho},r_{\rho}))  \\
  \text{two-hops} &= dist(H_{general}(n,c,\rho,d_{\rho},r_{\rho}),H_{general}(n,c+2,\rho,d_{\rho},r_{\rho}))  \\  
\end{align*}
Which are invarinat for all $n$ and $c$.

@@

By similar algebra and trigonometry:
\begin{align*}
  \text{two-hops}  &= \sqrt{\frac{4}{9} + r_{\rho}^2 (-\frac{16 (\sin^2( \rho / 2))}{9} + \sin^2(2\rho/3 + \frac{4\pi}{3})  + (1 - \cos(2\rho/3 + \frac{4\pi}{3}))^2)} \\
\end{align*}

We would really like to know the partial derivative of the two-hops - one-hop with respect
to the radius to be able to understand how to choose the radius to form the minimimax optimum.

@@

Then:
\begin{align*}
  \text{two-hops} - \text{one-hop}  &= \sqrt{\frac{4}{9}  + r_{\rho}^2(g_{\rho}+ j_{\rho})}
  - \sqrt{\frac{1}{9} +r_{\rho}^2(f_{\rho}+k_{\rho}) }
\end{align*}

% Using wolfram alph we se this derivate remains negative in this range....
% plot d/dr sin(2r/3 + 4pi/3)^2 - sin(r/3 + 2pi/3)^2 + (1 - cos(2r/3 + 4pi/3))^2 -(1-cos(r/3 + 2pi/3))^2 from 0 to 35.44 degrees
% in other words

% This part is also negative withing this range...
%  d/dr -(16/9)sin(r/2)^2 - -(4/9)sin(r/2)^2 from 0 to 35.44 degrees

%% Using Mathematica, I get:
%% f[x_, r_] := 
%%  Sqrt[4/9 + 
%%     x^2 (-16 Sin[r/2]^2/9 + 
%%        Sin[2 r/3 + 4 Pi /3]^2 + (1 - Cos[2 r/3 + 4 Pi/3])^2)] - 
%%   Sqrt[1/9 + 
%%     x^2*(-4*Sin[r/2]^2/9 + 
%%        Sin[r/3 + 2 Pi/3]^2 + (1 - Cos[r/3 + 2 Pi/3])^2) ]
%% N[N[f[1, 1]]]

%% Plot3D[D[f[x, r]], {x, 0, 2}, {r, 0, Pi}]

%% This graph demonstrates the derivative is always negative.

By graph inspection using Mathematica, we see the partial derivative of this with respect to
radius $r_{\rho}$ is always negative.
Since the partial derivative of $\text{two-hops} - \text{one-hop}$ with respect to the
radius is negative up until $\rho$ up until $\rho_{bc}$ when it is $0$, then
we will opitmize the overall minimax distance by choosing the largest radius
up until one-hop $= 1$, the rail-edge length. Numerical test suggests that this is true.

@@

\begin{align*}
  1 &=  \sqrt{\frac{1}{9}  + r_{opt}^2(-\frac{4 (\sin^2( \rho / 2))}{9} + \sin^2(\rho/3+ \frac{2\pi}{3})  + (1 - \cos(\rho/3 + \frac{2\pi}{3}))^2)} \\
  r_{opt} &= \frac{2}{\sqrt{\frac{9}{2} \cdot ( \sqrt{3} sin(\rho/3) + \cos(\rho/3)) + cos(\rho)+ 8 }} \numberthis  \label{eqrhoopt} \\
\end{align*}
(where denominator $  \neq 0$).

@@

\begin{align*}
  d_{opt}^2 &= 1 - 4 \frac{2}{\sqrt{\frac{9}{2} \cdot ( \sqrt{3} sin(\rho/3) + \cos(\rho/3)) + cos(\rho)+ 8 }}^2 (\sin{ \rho / 2})^2   \\
  d_{opt}^2 &= 1 - 4 \frac{4(\sin{ \rho / 2})^2}{\frac{9}{2} \cdot ( \sqrt{3} sin(\rho/3) + \cos(\rho/3)) + cos(\rho)+ 8 }    \\
  d_{opt}^2 &= 1 - \frac{16(\sin{ \rho / 2})^2}{9 ( \sqrt{3} sin(\rho/3) + \cos(\rho/3)) + cos(\rho)+ 8 }    \\
  d_{opt}^2 &= 1 - \frac{16 sin^2(\rho / 2) }{ cos(\rho) + 9( \sqrt{3} sin(\rho/3) + \cos(\rho/3)) + 8 }   \\
    d_{opt} &= \sqrt{1 - \frac{16 sin^2(\rho / 2) }{ cos(\rho) + 9( \sqrt{3} sin(\rho/3) + \cos(\rho/3)) + 8 }}    \numberthis  \label{dopt}  \\      
\end{align*}
Therefore, by computing $r_{opt}$ as a function of $\rho$ from this equation, we can construct minimax optimal tetrahelix for an $0 \leq \rho \leq \rho_{bc}$.

>>>>

\begin{abstract}
  The Boerdijk--Coxeter helix (BC helix, or tetrahelix) is a
  face-to-face stack of regular tetrahedra.  Considering the edges of
  these tetrahedra as structural members, the resulting structure is attractive and
  inherently rigid, and therefore interesting to architects,
  mechanical engineers, and robotocists.  A formula that matches the
  visually apparent helices forming the outer rails of the tetrahelix
  is derived which is convenient for designers.  This formula defines
  the vertices of tetrahelices of varying radius, pitch, and
  curvature, with the BC helix as a special case.  An addtional
  special case is of that of $0$ curvature or rail angle, which
  generates a \emph{tetrabeam}.  A particular choice of paramaters
  defines a novel object, the minimax optimal member length-difference
  tetrabeam, the \emph{equitetrabeam}.  A formula for computing the
  radius and z-axis travel that gives an optimal minimax tetrahelix
  for any rail angle between $0$ and $\rho_{bc}$ is given, thus
  defining a continuum of tetrahelices terminated by the equitetrabeam
  and BC helix. Numerically finding the rail angle from the equation for
  pitch allows optimal tetrahelices of any pitch to be designed.
  An interactive tool for such design and experimentation is provided.
  Utility and use for static and variable geometry
  truss/space frame design and robotics based on the are discussed.
\end{abstract}

@@

The Boerdijk--Coxeter helix\cite{coxeter1985simplicial} (BC helix), is
a face-to-face stack of tetrahedra that winds about a straight axis.
Because architects, structural engineers, and robotocists are inspired
by and follow such mathematical models but can build structures and
machines of differing or even dynamically changing length, it is
useful to develop the mathematics of structure formed from tetrahedra
where we relax regularity.

The vertices of the tetrahedra lie upon
three helices about the central axis.  The
Tetrobot/Glussbot\cite{TetrobotBook} project uses the regularity of
this geometry to make a tentacle-like robot that can crawl like a slug
or mollusc.  The Tetrobot concept is to use mechanical members, called
actuators, which can change their length, connected by special joints,
called the Song-Kwon-Kim\cite{song2003spherical} or turret joint,
which allow many members to come to a single point.  Such machines can
follow purely regular mathematical models such as the Boerdijk–-Coxeter
helix or the Octet Truss\cite{richard1961synergetic}.

@@

Imagining Figure \ref{fig:closeup} as a static mechanical structure,
we observe that it is useful to the mechanical engineer or
robotocist because the structure remains an inherently rigid,
omni-triangulated space frame, which may be expected to be at least
somewhat mechanically strong.
Imagine further in Figure \ref{fig:closeup}, that each static edge was replaced with an
actuator that could dynamically become shorter or longer in response to electronic control,
and the they vertices were a joint that supported sufficient angular displacement
for this to be possible. Such a machine is a Tetrobot\cite{TetrobotBook}.

BC helix does not rest on a plane in a simple way. It is convenient to
be able to ``untwist'' it and to form a tetrahelix space frame that has a
flat planar surface. By making length changes in a certain way, we can
untwist a tetrahelix to form a \emph{tetrabeam} which has planar faces
and has, for example, an equilateral triangular profile. This paper
develops the equations needed to untwist the tetrahelix. All math
deveoped here is available in JavaScript and demonstrated by an interactive
design website\cite{readtetrahelix}, from which Figures \ref{fig:closeup} and
\ref{fig:series} are taken.

@@

%% Note: This is not rendering figure two for some reason.
In Figure
\ref{fig:series}, the closest helix is the BC helix, and the furthest
is the equitetrabeam.

@@

If you can choose member lengths, you can form a linear combination of
the equitetrabream lengths and the completely regular lengths of the
tetrahelix, thereby choosing the amount of twist\footnote{The formal definiton
  of twistiness, or \emph{torsion}, is not useful or used in this paper, and
  we use \emph{curvature} without relying on the mathematical curvature easily
  computed from helical equations we provide.}.
If you are designing a
space frame, this is a static design choice, in a robot, it is a
dynamic choice that can be used to twist the robot and/or exert linear or
angular force on the environment.

Ideally we would have a simple formula for defining the nodes based on
any curvature or pitch we choose.  Unfortunately, it is not obvious that a linear
combination of lengths produces a simple formula.  It is a goal of
this paper to relate these two approaches to generating a tetrahelix
continuum.

@@

R.W. Gray's site\cite{graytetrahelix}, repeating a formula by Coxeter\cite{coxeter1985simplicial} in more accessible form, gives for a counter-clockwise BC Helix:
\begin{equation}
  \label{graycoxeter}
V(n) =
\left [
  \begin{tabular}{c}
   $ r_{bc} \cos(n \theta) $\\
   $ r_{bc} \sin(n \theta) $\\
   $ n h_{bc}  $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
 $ r_{bc} = \frac{3\sqrt{3}}{10} $\\
 $ h_{bc} = 1/\sqrt{10} $ \\
 $ \theta = \arccos(-2/3) $ \\
  \end{tabular}      
\end{equation}
where $n$ represents each integer numbered node in succession on every colored rail.

@@

Such a helix can be written:
\begin{equation}
  \label{eq:colord}
H_{BCcolored}(n,c) =
\left [
  \begin{tabular}{c}
   $ r_{bc}  \cos((3 \theta - 2 \pi)n + c  \theta $\\
   $ r_{bc} \sin((3 \theta - 2 \pi)n + c  \theta $\\
   $ 3 h_{bc} (n + c/3)  $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
 $ r_{bc} = \frac{3\sqrt{3}}{10} $\\
 $ h_{bc} = 1/\sqrt{10} $ \\
 $ \theta = \arccos(-2/3) $ \\
  \end{tabular}      
\end{equation}

@@

The quantity $ (3 \theta - 2 \pi) \approx 35.43 \degree $, and is the angular shift between $V(3n+c)=H_{BCcolored}(n,c)$ and
$V(3(n+1)+c)=H_{BCcolored}(n+1,c)$.
This quantity appears so often that we call it the ``rail angle rho''. For the BC helix, $\rho_{bc} = (3 \theta - 2 \pi)$.

@@

Note in Figure \ref{railanglefig} the $z$-axis travel for one rail edge is denoted by $d$. In \eqref{graycoxeter} and \label{eq:colored}, the variable
$h$ is used for one third of the distance we name $d$. We will later justify that $d = 3h$. In this paper we assume the length of a rail
is alwyas $1$ is a simplification, although we make the rail length a parameter in our JavaScript codein \texttt{tetrahelixmath.js} \cite{readtetrahelix}.

Since:
\[ \frac{2 \pi}{\rho_{bc}} \approx 10.16
\]
We can see that there are approximately $10.16$ red, blue or yellow tetrahedra on one rail in a single revolution.
The pitch of the Boerdijk--Coxeter helix of edge length $1$ is the length of three tetrahedra times this number:
\begin{align*}
  &= \frac{d_{bc} 2 \pi }{\rho_{bc}} \\  
  &= \frac{3 \cdot h_{bc} 2 \pi }{\rho_{bc}} \\
  &= \frac{3  \sqrt{\frac{2}{5}}  \pi}{\rho_{bc}} \\
  &\approx 9.64 \\
\end{align*}
The pitch is less than the number of tetrahedra because the tetrahedra
are not lined up perfectly.  It is a famous and interesting result
that the pitch is irrational, a BC helix never has two tetrahedra at
precisely the same orientation around the $z$-axis. However, this is
inconvenient to designers, who might prefer a rational pitch. For
example, a slight irregularity that led to a pitch of precisely 10
tetrahedra in one revolution would allow an architect to design a
column having a basis and a capital in the same relation to the
tetrahedra they touch.

@@

We use the term \emph{tetrahelix} to mean any structure made of
vertices and edges which is isomorphic to the BC helix, in which the
vertices lie on three helices, no matter what lengths the edges take
on. One could consider various definitions of optimality for a
tetrahelix, but the must useful to an enginner is to minimize the
maximum difference between any two edges. We call a tetrahelix
\emph{member-length minimax optimal} (or just \emph{optimal}) if there is not tetrahelix of the
same radius and pitch with a smaller maximum edge length difference.

@@

 Now that we have coincident axes, the same pitch, we can go on to
 the hard proof about where they occur.

@@

 In principle in any 3 helices with the same axis of the same radius
 having any relative
 displacement along the $z$ axis there are 9 distinct edge classes.
 If when projecting all vertices on the the $z$-axis, the interval
 defined by the $z$ axis value of its endpoints contains no other vertices,
 we call it a \emph{one-hop} edge, and if it does contain another vertex we
 call it a \emph{two-hop} edge.
 Then there are 
 3 rail edges, 3 one-hop lengths between each pair of 3 rails, and 3 two hop
 lengths between each pair of three rails, where the two-hop length is at least
 the one-hop length.
 However, we have already shown the rail
 lengths are equal in any optimal tetrahelix.

 \begin{theorem}
   \label{eventhirds}
   A tetrahelix of a given radius and axial distance $d$ in which all nodes are evenly spaced at $d/3$ intervals on the $z$ axis is optimal.
     Any one tetrahedron in a tetrahelix has $1$ rail edge, $2$ one-hop edges connected to the rail and $2$ two-hop edges connected to the rail.
  The edge opposite of the rail edge is a one-hop edge.
  \end{theorem}

 \begin{proof}
    Consider the tetrahelix in which the vertices are evenly spaced at
    $h/3$ intervals on the $z$ axis. Every edge is either a rail edge,
    or it makes one hop, or it makes two hops. All of the one-hop
    edges are equal length.  All of the two-hop edges are equal
    length.

@@

 Note that based on \ref{eventhirds}, there are only 3 possible lengths in an optimal tethrahelix,
 and we are justified in classifying edge lengths as \emph{rail},\emph{one-hop}, or
\emph{two-hop}. The one-hop edges are the edges between closest on the $z$-axis, and the two-hop edges are those that hop over a vertex.

By Theorem \ref{eventhirds} every optimal tetrahelix has vertices lying on helices expressible in the form:
\[
V_{optimal}(n,c) =
\left [
  \begin{tabular}{c}
   $ r \cos(n \alpha +  c 2 \pi /3)$\\
   $ r \sin(n \alpha +  c 2 \pi /3) $\\
   $ \frac{d(n +c / 3)}{3}   $
  \end{tabular}
  \right ],
\text{where:}
\begin{tabular}{c}
  $c \in \{0,1,2\}$
  \end{tabular}      
\]
where we have not yet investigated in the general case the relationships beteween $\alpha$, $r$, and $d$ in this formulation.
However, we understand that when $\alpha = 0$, the helices are degenerate, having curvature of $0$, and
we have the equitetrabeam.

This formulation $V(n,c)$ above is valuable, but obscures the essentially fact that the red, yellow, and blue helices distributed
about the central $z$ axis $120\degree$ from each other.
In order to rewrite this expression with an explicit rotation of $2\pi/3$, we expand 
the expression and seek to isolate the term $c2\pi/3$.
\begin{align*}
  \rho_{bc} n + c \theta  &=   \text{\{we aim for 3 in denominator, so we split...\}} \\
    (3 \theta - 2 \pi)n + (c/3)  (3 \theta)  &=   \text{\{we want $2\pi$ in numerator, so add canceling terms...\}} \\
  (3 \theta - 2 \pi)n + (c/ 3) (3 \theta - 2 \pi  + 2 \pi) &=  \text{\{associate...\}} \\
  (3 \theta - 2 \pi)n + (c/ 3) ((3 \theta - 2 \pi)  + 2 \pi) &=  \text{\{distribute...\}} \\  
  (3 \theta - 2 \pi)n + (c / 3) (3 \theta - 2 \pi)  + c 2 \pi /3 &=  \text{\{definition of $\rho_{bc}$...\}} \\
  \rho_{bc} n + (c / 3) \rho_{bc}  + c 2 \pi /3 &=  \text{\{collect like factors...\}} \\  
  \rho_{bc} (n + c/3)  + c 2 \pi /3  \\
\end{align*}
Now the the term on the left is the only one that depends on the scalar $n$. We use this to a create
a new formulation: $H_{BCsymmetric}(n,c) = H_{BCcolored}(n,c)$

@@

\section{Parameterizing Tetrahelices via Rail Angle}

We seek a formula to generate optimal tetrahelices that accepts a
parameter that allows us to design the tetrahelix conveniently.
Please refer back to Figure \ref{railanglefig}.
The pitch of the helix is an obious choice, but is not defined when the
curvature is $0$, an important special case. The radius or the axial
distance between two nodes on the same rail are obvious choices, but
perhaps the clearest choice is to build formula that takes as its
input the ``rail angle'' $\rho$. We define $\rho$ to be the angle
formed in the X,Y plane $\angle A O B$ projecting out the $z$
axis and sighting along the positive $z$ axis. In other words, $\rho$
controls how far a rail edge of a tetrahelix deviates from being
parallel with the axis, or the ``twistiness'' of tetrahelix. We use
the parameter $\chi = 1$ to indicate a chirality or counter-clockwise,
and $\chi = -1$ for clockwise.

@@

In choosing $\rho$, one greatly constrains $r$ and $d$, but does not completely
determine both of them together, so we treat both as parameters.

Rewriting our formulation in terms of $\rho$:
\begin{equation}
  \label{eq:general}
H_{general}(\chi,n,c,\rho,d_{\rho},r_{\rho}) =
\left [
  \begin{tabular}{c}
   $ r_{\rho} \cos(\chi \cdot (\rho \kappa + c 2 \pi /3)) $\\
   $ r_{\rho}  \sin(\chi \cdot (\rho \kappa + c 2 \pi /3)) $\\
   $ d_{\rho} \kappa $
  \end{tabular}
  \right ],
\text{where:}
\begin{tabular}{c}
  $   1 = d_{\rho}^2 + 4 r_{\rho}^2 (\sin{ \rho / 2})^2 $ \\
  $\kappa = n + c/3$ \\
    $\chi \in \{-1,1\}$ \\  
  \end{tabular}      
\end{equation}

@@

\begin{proof}
This is proved by a minimax argument.

By Theoerm \ref{eventhirds}, we can compute the (at most) three edge-lengths of an optimal
tetrahelix by (where $dist$ is the cartesian distance function):
\begin{align*}
  \text{rail} &= dist(H_{general}(n,c,\rho,d_{\rho},r_{\rho}),H_{general}(n+1,c,\rho,d_{\rho}),r_{\rho})) = 1 \\
  \text{one-hop} &= dist(H_{general}(n,c,\rho,d_{\rho},r_{\rho}),H_{general}(n,c+1,\rho,d_{\rho},r_{\rho}))  \\
  \text{two-hop} &= dist(H_{general}(n,c,\rho,d_{\rho},r_{\rho}),H_{general}(n,c+2,\rho,d_{\rho},r_{\rho}))  \\  
\end{align*}
Which are invarinat for all $n$ and $c$.

@@

By similar algebra and trigonometry:
\begin{align*}
  \text{two-hop}  &= \sqrt{\frac{4}{9} + r_{\rho}^2 (-\frac{16 (\sin^2( \rho / 2))}{9} + \sin^2(2\rho/3 + \frac{4\pi}{3})  + (1 - \cos(2\rho/3 + \frac{4\pi}{3}))^2)} \\
\end{align*}

We would really like to know the partial derivative of the two-hop - one-hop with respect
to the radius to be able to understand how to choose the radius to form the minimimax optimum.

@@

Then:
\begin{align*}
  \text{two-hop} - \text{one-hop}  &= \sqrt{\frac{4}{9}  + r_{\rho}^2(g_{\rho}+ j_{\rho})}
  - \sqrt{\frac{1}{9} +r_{\rho}^2(f_{\rho}+k_{\rho}) }
\end{align*}

By graph inspection using Mathematica, we see the partial derivative of this with respect to
radius $r_{\rho}$ is always negative.
Since the partial derivative of $\text{two-hop} - \text{one-hop}$ with respect to the
radius is negative up until $\rho$ up until $\rho_{bc}$ when it is $0$, then
we will opitmize the overall minimax distance by choosing the largest radius
up until one-hop $= 1$, the rail-edge length.

@@

\begin{align*}
  1 &=  \sqrt{\frac{1}{9}  + r_{opt}^2(-\frac{4 (\sin^2( \rho / 2))}{9} + \sin^2(\rho/3+ \frac{2\pi}{3})  + (1 - \cos(\rho/3 + \frac{2\pi}{3}))^2)} \\
  r_{opt} &= \frac{2}{\sqrt{\frac{9}{2} \cdot ( \sqrt{3} sin(\rho/3) + \cos(\rho/3)) + cos(\rho)+ 8 }} \numberthis  \label{eqrhoopt} \\
\end{align*}

@@

\begin{align*}
  d_{opt}^2 &= 1 - 4 \frac{2}{\sqrt{\frac{9}{2} \cdot ( \sqrt{3} sin(\rho/3) + \cos(\rho/3)) + cos(\rho)+ 8 }}^2 (\sin{ \rho / 2})^2   \\
  d_{opt}^2 &= 1 - 4 \frac{4(\sin{ \rho / 2})^2}{\frac{9}{2} \cdot ( \sqrt{3} sin(\rho/3) + \cos(\rho/3)) + cos(\rho)+ 8 }    \\
  d_{opt}^2 &= 1 - \frac{16(\sin{ \rho / 2})^2}{9 ( \sqrt{3} sin(\rho/3) + \cos(\rho/3)) + cos(\rho)+ 8 }    \\
  d_{opt}^2 &= 1 - \frac{16 sin^2(\rho / 2) }{ cos(\rho) + 9( \sqrt{3} sin(\rho/3) + \cos(\rho/3)) + 8 }   \\
    d_{opt} &= \sqrt{1 - \frac{16 sin^2(\rho / 2) }{ cos(\rho) + 9( \sqrt{3} sin(\rho/3) + \cos(\rho/3)) + 8 }}    \numberthis  \label{dopt}  \\      
\end{align*}

Thus, by computing $r_{opt}$  and $d_{opt}$ as a function of $\rho$ from this equation, we can construct minimax optimal tetrahelix for an $0 \leq \rho \leq \rho_{bc}$.
\end{proof}

\begin{theorem}
  \label{optimality}
  For any rail angle $\rho$ where $\rho \leq \rho_{bc}$, a tetrahelix of minimum maximum edge length
  difference is generated by $H_{general}(\chi,n,c,\rho,d_{\rho},r_{opt})$, where $r_{opt}$ and $d_{opt}$ are
  computed from $\rho$. 
\end{theorem}
